Parsing Postfix log files

### 2008-11-22 doc/logparser-acronyms.tex
Dealt with some XXX notes.
Switched to new glossaries package. A couple of other XXX items have been done too.

% % vim: set textwidth=75 spell :
% Warning: do not use acronyms within definitions, they do not work
% properly.

\newacronym{PLP}{Postfix Log Parser}{description={
    Postfix Log Parser is the parser implementing the algorithm described
    in this thesis.
}}

\newacronym{DNSBL}{DNS Blacklist}{description={
    A DNS Blacklist~\cite{Wikipedia-DNSBL} is a simple collaborative
    anti-spam technique used to reject or penalise email sent from mail
    servers reported to have sent large volumes of spam.  To use a DNSBL
    the mail server makes a DNS request incorporating the IP address of the
    client; if the requested hostname exists the client is on the DNSBL,
    and the mail server can decide what course of action to take.
}}

\newacronym{API}{Application Programming Interface}{description={
    One of the fundamental concepts when writing programs is the reuse of
    existing code, so that each program does not reinvent the wheel.  When
    a body of code is intended to be reused repeatedly, the user of this
    code needs to be informed of the functionality provided by the code.
    An API defines the interface provided to the user, and acts as a
    contract between the user and the provider: if the user adheres to the
    API the provider guarantees it will work, but is free to change the
    implementation provided the API is preserved.
}}

\newacronym{SMTP}{Simple Mail Transfer Protocol}{description={
    SMTP is the protocol that transfers mail from the sender to the
    recipient across the Internet.  It is a simple, human readable, plain
    text protocol, making it quite simple to test and debug problems with
    it.  A detailed description of \SMTP{} is beyond the scope of this
    document: the original protocol definition is \RFC{}~821~\cite{RFC821},
    updated in \RFC{}~2821~\cite{RFC2821}; introductory guides can be found
    at~\cite{smtp-intro-01,smtp-intro-02}.
}}

\newacronym{LMTP}{Local Mail Transfer Protocol}{description={
    LMTP is a protocol derived from SMTP that removes the need for the
    server to maintain a mail delivery queue, instead relying on the client
    to maintain it.  Typically the client is an MTA, and the server a
    delivery agent or a mail store.  Full details are available
    in~\cite{lmtp-rfc}.
}}

\newacronym{MTA}{Mail Transfer Agent}{description={
    A Mail Transfer Agent sends and receives mail via
    SMTP\backslashat{}.  XXX DO I NEED THIS\backslashat{}?
}}

\newacronym{RFC}{Request For Comments}{description={
    The Request For Comments series is a series of proposals defining
    various protocols and file formats, e.g.\ SMTP\backslashat{}.  The name
    is somewhat misleading: initially the authors were asking for peer
    review, but these documents are now the de facto standards the Internet
    runs on.
}}

\newacronym{SQL}{Structured Query Language}{description={
    SQL is the standard language used for database querying, modification
    and maintenance.  Some information about SQL, including its history,
    can be found at \url{http://en.wikipedia.org/wiki/SQL}; a good
    introduction can be found at \url{http://philip.greenspun.com/sql/}.
}}

\newacronym{LMA}{Log Mail Analyzer}{description={
    One of the other Postfix log parsers reviewed and the only example of
    prior published paper~\cite{log-mail-analyser} found that specifically
    deals with parsing Postfix log files.
}}

\newacronym{CSV}{Comma-Separated Value}{description={
    The most basic form of database available, each record is a single line
    in the file and the fields are separated by a special character,
    typically a comma or colon.
}}

\newacronym{FQDN}{Fully Qualified Domain Name}{description={
    An FQDN is a hostname plus domain name, e.g.\ \newline{}
        example.com     is a domain name          \newline{}
        www             is a hostname             \newline{}
        www.example.com is a FQDN
}}

\newacronym{pid}{Process Identifier}{description={
    There may be multiple copies of any program executing at any one time,
    so the program's name is not suitable as a distinguishing identifier;
    instead each process executing is given a pid that is guaranteed to be
    unique for the lifetime of the process.  Once the process has
    completed, the pid may be reused, as they are drawn from a finite pool.
}}

\newacronym{UCE}{Unsolicited Commercial Email}{description={
    UCE is a more restrictive definition of spam than most people use: it
    only covers mail that is explicitly commercial, thus excluding viruses,
    Bayesian poisoning mails, backscatter, and those annoying chain letters
    you get from friends.
}}

% I think this is better than having both regex and regexes in the acronym
% list.  Likewise for pids.
\newcommand{\pids}{\pid{}s}

\newacronym{ESMTP}{Extended SMTP}{description={
    ESMTP is Extended SMTP, defined in RFC~1869~\cite{RFC1869}.
}}

\newacronym{IP}{Internet Protocol}{description={
    The Internet Protocol is the protocol used to communicate between
    computers on the Internet.  An IP address is a unique address assigned
    to a computer on the Internet, allowing it to communicate with other
    computers on the Internet.
}}

\newacronym{SLCT}{Simple Logfile Clustering Tool}{description={
    SLCT~\cite{slct-paper} is a tool implementing an algorithm designed by
    Risto Vaarandi for identifying, grouping and producing a regex to match
    similar log lines.
}}

\newacronym{ATN}{Augmented Transition Networks}{description={
    Augmented Transition Networks, originally described in~\cite{atns} and
    further in~\cite{nlpip}, are a tool used in Computational Linguistics
    for creating grammars to parse or generate sentences,
}}

\newacronym{CLI}{Command Line Interface}{description={
    An computer interface based on typing commands, rather than using a
    mouse.
}}

\newacronym{NLP}{Natural Language Processing}{description={
    Natural Language Processing~\cite{Wikipedia-nlp} attempts to increase
    our understanding of the languages normally used by humans (e.g.\
    English, Spanish, Japanese), with the goal of writing programs that
    can understand human languages.
}}

>>>>

% % vim: set textwidth=75 spell :
% Note: every acronym is listed twice: once so it can be expanded, and once
% so it will appear with a long description in the glossary.
% Warning: do not use acronyms within definitions, they do not work
% properly.

\newglossaryentry{PLP}{
    type=\acronymtype,
    name={PLP},
    description={Postfix Log Parser\glsadd{PLP glossary}}
}
\newglossaryentry{PLP glossary}{
    name={Postfix Log Parser},
    description={
        Postfix Log Parser is the parser implementing the algorithm
        described in this thesis.
    }
}

\newglossaryentry{DNSBL}{
    type=\acronymtype,
    name={DNSBL},
    description={DNS Blacklist\glsadd{DNSBL glossary}}
}
\newglossaryentry{DNSBL glossary}{
    name={DNS Blacklist},
    description={
        A DNS Blacklist~\cite{Wikipedia-DNSBL} is a simple collaborative
        anti-spam technique used to reject or penalise email sent from mail
        servers reported to have sent large volumes of spam.  To use a
        DNSBL the mail server makes a DNS request incorporating the IP
        address of the client; if the requested hostname exists the client
        is on the DNSBL, and the mail server can decide what course of
        action to take.
    }
}

\newglossaryentry{API}{
    type=\acronymtype,
    name={API},
    description={Application Programming Interface\glsadd{API glossary}}
}
\newglossaryentry{API glossary}{
    name={Application Programming Interface},
    description={
        One of the fundamental concepts when writing programs is the reuse
        of existing code, so that each program does not reinvent the wheel.
        When a body of code is intended to be reused repeatedly, the user
        of this code needs to be informed of the functionality provided by
        the code.  An API defines the interface provided to the user, and
        acts as a contract between the user and the provider: if the user
        adheres to the API the provider guarantees it will work, but is
        free to change the implementation provided the API is preserved.
    }
}

\newglossaryentry{SMTP}{
    type=\acronymtype,
    name={SMTP},
    description={Simple Mail Transfer Protocol\glsadd{SMTP glossary}}
}
\newglossaryentry{SMTP glossary}{
    name={Simple Mail Transfer Protocol},
    description={
        SMTP is the protocol that transfers mail from the sender to the
        recipient across the Internet.  It is a simple, human readable,
        plain text protocol, making it quite simple to test and debug
        problems with it.  A detailed description of SMTP is beyond the
        scope of this document: the original protocol definition is
        RFC~821~\cite{RFC821}, updated in RFC~2821~\cite{RFC2821};
        introductory guides can be found
        at~\cite{smtp-intro-01,smtp-intro-02}.
    }
}

\newglossaryentry{LMTP}{
    type=\acronymtype,
    name={LMTP},
    description={Local Mail Transfer Protocol\glsadd{LMTP glossary}}
}
\newglossaryentry{LMTP glossary}{
    name={Local Mail Transfer Protocol},
    description={
        LMTP is a protocol derived from SMTP that removes the need for the
        server to maintain a mail delivery queue, instead relying on the
        client to maintain it.  Typically the client is an MTA, and the
        server a delivery agent or a mail store.  Full details are
        available in~\cite{lmtp-rfc}.
    }
}

\newglossaryentry{MTA}{
    type=\acronymtype,
    name={MTA},
    description={Mail Transfer Agent\glsadd{MTA glossary}}
}
\newglossaryentry{MTA glossary}{
    name={Mail Transfer Agent},
    description={
        A Mail Transfer Agent sends and receives mail via
        SMTP\backslashat{}.  Users submit mail to an MTA via their mail
        client (e.g.\ Microsoft Outlook, Thunderbird, webmail services);
        the sending MTA transfers the mail to the receiving MTA (which may
        in turn forward it to another MTA).
    }
}

\newglossaryentry{RFC}{
    type=\acronymtype,
    name={RFC},
    description={Request For Comments\glsadd{RFC glossary}}
}
\newglossaryentry{RFC glossary}{
    name={Request For Comments},
    description={
        The Request For Comments series is a series of proposals defining
        various protocols and file formats, e.g.\ SMTP\backslashat{}.  The
        name is somewhat misleading: initially the authors were asking for
        peer review, but these documents are now the de facto standards the
        Internet runs on.
    }
}

\newglossaryentry{SQL}{
    type=\acronymtype,
    name={SQL},
    description={Structured Query Language\glsadd{SQL glossary}}
}
\newglossaryentry{SQL glossary}{
    name={Structured Query Language},
    description={
        SQL is the standard language used for database querying,
        modification and maintenance.  Some information about SQL,
        including its history, can be found at
        \url{http://en.wikipedia.org/wiki/SQL}; a good introduction can be
        found at \url{http://philip.greenspun.com/sql/}.
    }
}

\newglossaryentry{LMA}{
    type=\acronymtype,
    name={LMA},
    description={Log Mail Analyzer\glsadd{LMA glossary}}
}
\newglossaryentry{LMA glossary}{
    name={Log Mail Analyzer},
    description={
        One of the other Postfix log file parsers reviewed and the only
        example of prior published paper~\cite{log-mail-analyser} found
        that specifically deals with parsing Postfix log files.
    }
}

\newglossaryentry{CSV}{
    type=\acronymtype,
    name={CSV},
    description={Comma-Separated Value\glsadd{CSV glossary}}
}
\newglossaryentry{CSV glossary}{
    name={Comma-Separated Value},
    description={
        The most basic form of database available, each record is a single
        line in the file and the fields are separated by a special
        character, typically a comma or colon.
    }
}

\newglossaryentry{FQDN}{
    type=\acronymtype,
    name={FQDN},
    description={Fully Qualified Domain Name\glsadd{FQDN glossary}}
}
\newglossaryentry{FQDN glossary}{
    name={Fully Qualified Domain Name},
    description={
        An FQDN is a hostname plus domain name, e.g.\ \newline{}
            example.com     is a domain name          \newline{}
            www             is a hostname             \newline{}
            www.example.com is a FQDN
    }
}

\newglossaryentry{pid}{
    type=\acronymtype,
    name={pid},
    description={Process Identifier\glsadd{pid glossary}}
}
\newglossaryentry{pid glossary}{
    name={Process Identifier},
    description={
        There may be multiple copies of any program executing at any one
        time, so the program's name is not suitable as a distinguishing
        identifier; instead each process executing is given a pid that is
        guaranteed to be unique for the lifetime of the process.  Once the
        process has completed, the pid may be reused, as they are drawn
        from a finite pool.
    }
}

\newglossaryentry{UCE}{
    type=\acronymtype,
    name={UCE},
    description={Unsolicited Commercial Email\glsadd{UCE glossary}}
}
\newglossaryentry{UCE glossary}{
    name={Unsolicited Commercial Email},
    description={
        UCE is a more restrictive definition of spam than most people use:
        it only covers mail that is explicitly commercial, thus excluding
        viruses, Bayesian poisoning mails, backscatter, and those annoying
        chain letters you get from friends.
    }
}

\newglossaryentry{ESMTP}{
    type=\acronymtype,
    name={ESMTP},
    description={Extended SMTP\glsadd{ESMTP glossary}}
}
\newglossaryentry{ESMTP glossary}{
    name={Extended SMTP},
    description={
        ESMTP is Extended SMTP, defined in RFC~1869~\cite{RFC1869}.
    }
}

\newglossaryentry{IP}{
    type=\acronymtype,
    name={IP},
    description={Internet Protocol\glsadd{IP glossary}}
}
\newglossaryentry{IP glossary}{
    name={Internet Protocol},
    description={
        The Internet Protocol is the protocol used to communicate between
        computers on the Internet.  An IP address is a unique address
        assigned to a computer on the Internet, allowing it to communicate
        with other computers on the Internet.
    }
}

\newglossaryentry{SLCT}{
    type=\acronymtype,
    name={SLCT},
    description={Simple Logfile Clustering Tool\glsadd{SLCT glossary}}
}
\newglossaryentry{SLCT glossary}{
    name={Simple Logfile Clustering Tool},
    description={
        SLCT~\cite{slct-paper} is a tool implementing an algorithm designed
        by Risto Vaarandi for identifying, grouping and producing a regex
        to match similar log lines.
    }
}

\newglossaryentry{ATN}{
    type=\acronymtype,
    name={ATN},
    description={Augmented Transition Networks\glsadd{ATN glossary}}
}
\newglossaryentry{ATN glossary}{
    name={Augmented Transition Networks},
    description={
        Originally described in~\cite{atns} and further in~\cite{nlpip},
        Augmented Transition Networks are a tool used in Computational
        Linguistics for creating grammars to parse or generate sentences,
    }
}

\newglossaryentry{CLI}{
    type=\acronymtype,
    name={CLI},
    description={Command Line Interface\glsadd{CLI glossary}}
}
\newglossaryentry{CLI glossary}{
    name={Command Line Interface},
    description={
        An computer interface based on typing commands, rather than using a
        mouse.
    }
}

\newglossaryentry{NLP}{
    type=\acronymtype,
    name={NLP},
    description={Natural Language Processing\glsadd{NLP glossary}}
}
\newglossaryentry{NLP glossary}{
    name={Natural Language Processing},
    description={
        Natural Language Processing~\cite{Wikipedia-nlp} attempts to
        increase our understanding of the languages normally used by humans
        (e.g.\ English, Spanish, Japanese), with the goal of writing
        programs that can understand human languages.
    }
}

### 2008-11-22 doc/logparser-architecture.tex
Switched to new glossaries package. A couple of other XXX items have been done too.

There is some similarity between the parser's design and William Wood's
\ATN{}~\cite{atns,nlpip}, a tool used in Computational Linguistics for
creating grammars to parse or generate sentences.  The resemblance between
\ATN{} and the parser is accidental, but it is obvious that the two
different approaches share an underlying separation of concerns; this
appears to be a natural division of responsibility and functionality.
\Tableref{Similarities between ATN and this architecture} shows the
similarities.

@@

\begin{table}[ht]
    \caption{Similarities between ATN and this architecture}
    \empty{}\label{Similarities between ATN and this architecture}
    \begin{tabular}[]{lll}
        \tableline{}
        \ATN{}          & Architecture  & Similarity                \\
        \tableline{}
        Networks        & Algorithm     & Determines the sequence 
                                        of transitions              \\
                        &               & or actions that 
                                        constitutes a valid  input. \\
        Transitions     & Actions       & Save data and impose
                                        conditions the              \\
                        &               & input must meet to be
                                        accepted as valid.          \\
        Abbreviations   & Rules         & Responsible for 
                                        classifying input.          \\
        \tableline{}
    \end{tabular}
\end{table}

@@

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.  Unintentionally overlapping rules lead to
inconsistent parsing and data extraction because the order in which rules
are tried against each line may change between log files, and the first
matching rule wins.  Overlapping rules are frequently a requirement,
allowing a more specific rule to match some log lines and a more general
rule to match the majority, e.g.\ separating \SMTP{} delivery to specific
sites from \SMTP{} delivery to the rest of the world.  The algorithm
provides a mechanism for ordering overlapping rules: the priority field in
each rule (defaults to zero).  Negative priorities may be useful for
catchall rules.

@@

    \item Sort by regex and visually inspect the list, e.g.\ with \SQL{}
        similar to: \textbf{select regex from rules order by regex;}

>>>>

There is some similarity between the parser's design and William Wood's
\gls{ATN}~\cite{atns,nlpip}, a tool used in Computational Linguistics for
creating grammars to parse or generate sentences.  The resemblance between
\gls{ATN} and the parser is accidental, but it is obvious that the two
different approaches share an underlying separation of concerns; this
appears to be a natural division of responsibility and functionality.
\Tableref{Similarities between ATN and this architecture} shows the
similarities.

@@

\begin{table}[ht]
    \caption{Similarities between ATN and this architecture}
    \empty{}\label{Similarities between ATN and this architecture}
    \begin{tabular}[]{lll}
        \tableline{}
        \gls{ATN}          & Architecture  & Similarity                \\
        \tableline{}
        Networks        & Algorithm     & Determines the sequence 
                                        of transitions              \\
                        &               & or actions that 
                                        constitutes a valid  input. \\
        Transitions     & Actions       & Save data and impose
                                        conditions the              \\
                        &               & input must meet to be
                                        accepted as valid.          \\
        Abbreviations   & Rules         & Responsible for 
                                        classifying input.          \\
        \tableline{}
    \end{tabular}
\end{table}

@@

The parser does not try to detect overlapping rules; that responsibility is
left to the author of the rules.  Unintentionally overlapping rules lead to
inconsistent parsing and data extraction because the order in which rules
are tried against each line may change between log files, and the first
matching rule wins.  Overlapping rules are frequently a requirement,
allowing a more specific rule to match some log lines and a more general
rule to match the majority, e.g.\ separating \gls{SMTP} delivery to
specific sites from \gls{SMTP} delivery to the rest of the world.  The
algorithm provides a mechanism for ordering overlapping rules: the priority
field in each rule (defaults to zero).  Negative priorities may be useful
for catchall rules.

@@

    \item Sort by regex and visually inspect the list, e.g.\ with \gls{SQL}
        similar to: \textbf{select regex from rules order by regex;}
